Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-02-08 ThesisJP/Background/background.tex
Small corrections.

\section{General-purpose computing on graphics processing units}\label{sec:gpgpu}
In a recent essay, Herb Sutter, a leading authority on software development, reviews the current state of the software and computer industry\todo{cite}. The essay bears the name "Welcome to the Jungle" and is a sequel to the essay "The Free Lunch is Over" from 2004. The two essays spins around the fact that chip manufactures hit a frequency-wall in the beginning of the 21st century. Until then, all mainstream computer programs where typically running in a single thread on a single core, and a programmer could expect the software to annually gain performance without touching the code. This was the "free-lunch" era as depicted in Fig.\,\ref{fig:jungle}. The increase in processing power came mostly from an ever-growing clock frequency. However, power consumption and heat generation where also growing with the increase in clock frequency, until the level of cooling required finally became to much in 2004.  From that point, chip manufactures had to pursue a different approach  than just increasing the clock frequency. The immediate solution was to embed several cores in one CPU chip. This insured continuation of Moore's law, since the number of transistors per chip could continue to grow. For programmers, multi-core CPUs meant that computer programs now had to be multi-threaded in order to continue to gain performance. It was a concurrency revolution,and  the free lunch was over.

@@

Around the same time as CPU manufactures hit the frequency wall, people started experimented with programmable shaders which recently had been introduced in the field of graphics programming. The graphics processing pipeline had previously been a fixed function pipeline, where e.g. vertices and texture coordinates where fed to the graphics processing unit (GPU) in one end and a rasterized rendered image was outputted in the other. What programmable shaders  introduced was the ability to replace some steps in the fixed-function pipeline with custom computations. This made it possible to utilize the GPU for other tasks than just rendering graphics. The graphics pipeline could for instance be hacked to evaluate a differential equation over grid defined by graphics primitives\todo{cite Seland et al.}. The rationale behind this exploration was that hidden in the GPU was processing power an order of magnitude larger than what the CPU could provide at the same time. Actually, at this time, the GPU had already gone multi-core, driven by the increasing demand for realistic computer games. Multiple cores was somewhat easier to achieve with the GPU since the rendering problem is inherently parallel. The pixels in the rendered image can be calculated in any order, and one pixel is therefore an concealed unit which the GPU scheduler is free to issue at any time to whatever core that is available. It it is also much less costly to switch between a pixel-thread for the GPU then a program thread for the CPU. Since there was only one program running on the GPU, the fixed function pipeline, GPU designers had a good idea of how the program would execute, and could therefore skip advance feature as predicting execution paths and different levels of data caching.  The released silicon was used to make more cores.

>>>>

\section{General-purpose computing on graphics processing units}\label{sec:gpgpu}
In a recent essay, Herb Sutter, a leading authority on software development, reviews the current state of the software and computer industry\todo{cite}. The essay bears the name "Welcome to the Jungle" and is a sequel to the essay "The Free Lunch is Over" from 2004. The two essays spins around the fact that chip manufactures hit a frequency-wall in the beginning of the 21st century. Until then, all mainstream computer programs where typically running in a single thread on a single core, and a programmer could expect the software to annually gain performance without touching the code. This was the "free-lunch" era as depicted in Fig.\,\ref{fig:jungle}. The increase in processing power came mostly from an ever-growing clock frequency. However, power consumption and heat generation where also growing at the same rate, until the level of cooling required finally became to much in 2004.  From that point, chip manufactures had to pursue a different approach  than just increasing the clock frequency. The immediate solution was to embed several cores in one CPU chip. This insured continuation of Moore's law, since the number of transistors per chip could continue to grow. For programmers, multi-core CPUs meant that computer programs now had to be multi-threaded in order to continue to gain performance. It was a concurrency revolution, and  the free lunch was over.

@@

Around the same time as CPU manufactures hit the frequency wall people started experimented with programmable shaders which recently had been introduced in the field of graphics programming. The graphics processing pipeline had previously been a fixed function pipeline, where e.g. vertices and texture coordinates where fed to the graphics processing unit (GPU) in one end and a rasterized rendered image was outputted in the other. What programmable shaders  introduced was the ability to replace some steps in the fixed-function pipeline with custom computations. This made it possible to utilize the GPU for other tasks than just rendering graphics. The graphics pipeline could for instance be hacked to evaluate a differential equation over grid defined by graphics primitives\todo{cite Seland et al.}. The rationale behind this exploration was that hidden in the GPU was processing power an order of magnitude larger than what the CPU could provide at the same time. Actually, at this time, the GPU had already gone multi-core, driven by the increasing demand for realistic computer games. Multiple cores was somewhat easier to achieve with the GPU since the rendering problem is inherently parallel. The pixels in the rendered image can be calculated in any order, and one pixel is therefore an concealed unit which the GPU scheduler is free to issue at any time to whatever core that is available. It it is also much less costly to switch between a pixel-thread for the GPU then a program thread for the CPU. Since there was only one program running on the GPU, the fixed function pipeline, GPU designers had a good idea of how the program would execute, and could therefore skip advance feature as predicting execution paths and different levels of data caching.  The released silicon was used to make more cores.

### 2014-02-08 ThesisJP/Introduction/introduction.tex
Small corrections.

\section{Motivation}
A rapid development in computer game technology and accompanying programming languages have recently provided researchers with small personal super computers, comprised in a single graphics card (\nom{GPU}{Graphics processing unit}). The latest graphics cards\footnote{As of January 2014.} from both \nom{Nvidia}{Graphics card producer} and \nom{AMD}{Graphics card producer} provide close to 6 T\nom{FLOPS}{Floating point operations per second} single precision and 1.5 TFLOPS double precision in computational power. For double precision this is more than the worlds most powerful supercomputer could provide in 1998.

>>>>

\section{Motivation}
A rapid development in computer game technology and accompanying programming languages have recently provided researchers with small personal super computers, comprised in a single graphics card (\nom{GPU}{Graphics processing unit}). The latest graphics cards\footnote{As of January 2014.} from both \nom{Nvidia}{Graphics card producer} and \nom{AMD}{Graphics card producer} provide close to 6 T\nom{FLOPS}{Floating point operations per second} single precision and 1.5 TFLOPS double precision computations. For double precision this is more than the worlds most powerful supercomputer could provide in 1998.
